Single Bivalent Level Analysis

### 2020-01-24 doc/Bivalent_Analysis.Rmd
some work from tonight

@@

#This isn't the same as curated/passing bivalents           
#This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
#Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally #segmented.

@@

@@

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")
show.the.cells_females_dom
#PWD seems to have the biggest variance across cells

almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+

  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_DOM_f

@@

female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")
female_long_biv_plot

male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")
male_long_biv_plot

@@

#code if males are high or low?

@@

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)) ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")

@@

ToDo fix the function to calculat IFD2, IFD3

@@

How can I highlight the result (I need to verify this first) for greater IFD in high rec male mice?

@@

```{r show.IFD, echo=FALSE, warning=FALSE}
colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                    "F1"='grey', "other" = 'grey'
                    )
#first draft of the IFD comparisons
#make the mouse boxplots -- instead of geom_points

@@

#boxplots
boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")
boxplots_first.IFD.plot_ABS
#order is out of wack, but -- good for now

boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))
boxplots_first.IFD.plot_PER

first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")

@@

#+theme(legend.position="none")

```{r blah.again}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData

#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")

long.biv_IFD_look_ABS

long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")

long.biv_IFD_look_PER

Above is the first plots of IFDs. Each point is a IFD observation. colors are by mouse. 
Maybe add mouse level lines for the means

Maybe I should try binned the SC lengths.

mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.

@@

```{r MM_1.SC.Length, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

@@

### second set of Mixed Models

@@

library(RLRsim)
IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)

anova(IFD_ABS.MM)

#thing1  <- summary(IFD_ABS.MM)
IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq") 
IFD.ABS_chi

@@

IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)

anova(IFD_PER.MM)

Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
Chis.sp_IFD.PER.MM

@@

```{r compare.IFD}

#make t.tests for pooled high and low

@@

@@

@@

for the long chrm, pull the object ids out of the melt table.

@@

CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")
CO.Pos.density_1CO

@@

Above plot focuses on the 1CO bivalent normalized positions -- since CO interference controls the general position of COs when there are multiple COs. This plot shows the sexual dimorphism in the density plots.

consider adding annotate_text for the number of observations in each category.
think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.

@@

Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")
Dom.male_MELT_long

@@

Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")
Musc_male_MELT_long

#there are v few observations in this dataset

Tired runnin plots for the long biv data set, I think there are too few observations.

add boxplots for

@@

CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("boxy plots")+geom_jitter(alpha=.1)
CO.Pos.boxplots_1CO

@@

@@

#library(lmerTest)
#anova(CO_position.MM)
anova(CO_position.MM, verbose = TRUE)

summary(CO_position.MM) #no p values,,
#these just give t and F stats, no pvalue

@@

#test high vs low Musc strains
mouse.avs_4MM

@@

# old drafts

@@

#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
```

>>>>

```{r misc.setup, echo=FALSE}

colors_of_strains <- c('WSB'= "#56B4E9",'G' ='cadetblue','LEW'= 'lightblue','PERC'= 'blue',
                       #blues

                    "F1"='grey', "other" = 'grey'
                    )
```

@@

#This isn't the same as curated/passing bivalents           
#This plot illustrates how well the algorithm seperated indiviual bivalents per cell image. 
#Several MSM male spreads had tight cells with a high degree of overlap, so fewer bivalents were indiviudally #segmented.
```

@@

#colors_of_strains
#scale_color_manual(values=colors_of_strains)

@@

  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Female whole cell manual measures")

show.the.cells_females_dom
#PWD seems to have the biggest variance across cells

almost_cells_DOM_f <- ggplot(almost_whole_cell_BivData[(almost_whole_cell_BivData$sex == "female") & (almost_whole_cell_BivData$subsp == "Dom"),], aes(x=fileName, y= chromosomeLength, color=strain))+
  geom_point()+ 
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=mean_SC, x=fileName, col="red", size=2))+
  geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=median_SC, x=fileName, col="black", size=1.5))+
   geom_point(data=Biv_Data_mean_table[(Biv_Data_mean_table$sex == "female") & (Biv_Data_mean_table$subsp == "Dom"),],aes(y=Qar.4, x=fileName, col="black", size=2.6), shape=4)+
  scale_color_manual(values=colors_of_strains)+
  facet_wrap(~strain, scales="free")+ylim(c(30,200))+ggtitle("Dom female, almost whole cell \nMean, median, 4th Quartille ploted")
almost_cells_DOM_f

@@

female_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "female",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nFemale")+scale_color_manual(values=colors_of_strains)
female_long_biv_plot

male_long_biv_plot <- ggplot(real.long.bivData[real.long.bivData$sex == "male",], aes(x=fileName, y=chromosomeLength, color=strain))+geom_point()+facet_wrap(~category, scales="free")+ylim(c(30,200))+ggtitle("Long Bivalent Dataset, \nmale")+scale_color_manual(values=colors_of_strains)
male_long_biv_plot

@@

#code if males are high or low?
```

@@

first.SC.plot <- ggplot(Curated_BivData, aes(y = chromosomeLength, x= strain, fill= hand.foci.count))+geom_jitter(aes(alpha=0.2))+geom_boxplot(aes(alpha=0.5))+  ggtitle("SC Length for Bivalents")+facet_wrap(sex~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)
first.SC.plot
#add boxplots for mouse level means
#add second x axis, hand.foci.count

other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)

@@

3) Interference / IFD will not be sexually dimorphic.

B.1) **Interfernce/IFD** will be shorter in high Rec strains to allow more foci to fit on a single bivalent.

@@

@@

 Each point is a IFD observation. Maybe add mouse level lines for the means.

```{r show.IFD, echo=FALSE, warning=FALSE}

#first draft of the IFD comparisons
#make the mouse boxplots -- instead of geom_points

@@

#boxplots
boxplots_first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
boxplots_first.IFD.plot_ABS
#order is out of wack, but -- good for now

boxplots_first.IFD.plot_PER <- ggplot(Curated_BivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, PER")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+ylim(c(0,1))+scale_color_manual(values=colors_of_strains)
boxplots_first.IFD.plot_PER

first.IFD.plot_ABS <- ggplot(Curated_BivData.22, aes(y = IFD1_ABS, x= strain, color=mouse))+geom_jitter()+ggtitle("Interfocal Distances, ABS")+facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none") #+scale_color_manual(values=colors_of_strains)

@@

#+theme(legend.position="none")
```

```{r blah.again, echo=FALSE, eval=FALSE}
#long SC Long_biv_mouse.avs_4MM
#real.long.bivData
#Only use 2CO bivalents
real.long.bivData.22 <- real.long.bivData[real.long.bivData$hand.foci.count == 2,]
real.long.bivData.22 <- real.long.bivData.22[(!is.na(real.long.bivData.22$hand.foci.count)),]

long.biv_IFD_look_ABS <- ggplot(real.long.bivData.22, aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(20,160))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

long.biv_IFD_look_PER <- ggplot(real.long.bivData.22, aes(y = IFD1_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+  facet_wrap(sex~subsp, scales = "free")+ylim(c(0,1))+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)
```

### Fixed Effects

Mixed model analysis for IFD (interference), the first set of models are made with the lme() functions.

@@

```{r MM_IFB.ABS, echo=FALSE}
#get a list of the old mice to double check
#old.mice <- DF.HetC.MixedModel.HQ$mouse[DF.HetC.MixedModel.HQ$age.weeks < 25,]
#MIXED MODEL

@@

Sex is a slightly significant factor for raw and normalized IFD.

### Random Effects

@@

IFD_ABS.MM <- lmer(mean_IFD.2CO_ABS ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(IFD_ABS.MM)

#thing1  <- summary(IFD_ABS.MM)
IFD.ABS_chi <- drop1(IFD_ABS.MM, test="Chisq")

@@

IFD_PER.MM <- lmer(mean_IFD.2CO_PER ~ subsp * sex + (1|strain), data=mouse.avs_4MM, REML = TRUE)
#anova(IFD_PER.MM)

Chis.sp_IFD.PER.MM <- drop1(IFD_PER.MM, test="Chisq") 
#Chis.sp_IFD.PER.MM

@@

Random effect of strain is not significnt for ABS IFD, and only slightly significant for the IFD.PER

#### IFD 3CO bivalents

Run comparisons for 3CO bivalents.

@@

```{r compare.IFD,echo=FALSE}
#make t.tests for pooled high and low
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM$strain), 1, 0))

Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                    ifelse(grepl("MSM male", Curated_BivData$category), 1, 0))

Curated_BivData <- add_IFD(Curated_BivData)

Rec.group_IFD.ABS_test <- t.test(Curated_BivData$IFD1_ABS[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
     Curated_BivData$IFD1_ABS[ ( Curated_BivData$Rec.group ==1 )]
)#p-value = 7.955e-05

Rec.group_IFD.ABS_test$p.value

Rec.group_IFD.PER_test <- t.test( Curated_BivData$IFD1_PER[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc") & (Curated_BivData$Rec.group ==0 )],
     Curated_BivData$IFD1_PER[ ( Curated_BivData$Rec.group ==1 )]
)#p-value = 0.04231

Rec.group_IFD.PER_test$p.value

biv.IFDs <- ggplot(Curated_BivData[ ( Curated_BivData$sex == "male" ) & (Curated_BivData$subsp == "Musc"),], aes(y = IFD1_ABS, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~Rec.group, scales = "free")+theme(legend.position="none")+ylim(c(0,100))

oo.ifd <- ggplot(mouse.avs_4MM[ ( mouse.avs_4MM$sex == "male" ) & (mouse.avs_4MM$subsp == "Musc"),], aes(y = mean_IFD.2CO_PER, x= mouse, color=strain))+geom_boxplot(alpha=.2)+
  geom_jitter(alpha=.2)+
  ggtitle("Interfocal Distances, ABS")+
  scale_color_manual(values=colors_of_strains)+ facet_wrap(~strain, scales = "free")+ylim(c(0,1))+theme(legend.position="none")

@@

The t.tests for IFD1s at the bivalent level for the high and low musc males are significant.

@@

#### Predictions

**Heterochiasmy**

**Polymorphism**

@@

CO.Pos.density_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(sex~subsp, ncol=2)+ggtitle("Normalized 1CO Foci Positions")+scale_fill_manual(values=colors_of_strains)
CO.Pos.density_1CO

@@

Above plot focuses on the 1CO bivalent normalized positions since CO interference controls the general position of COs when there are multiple COs. This plot shows the sexual dimorphism in the density plots.

Consider adding annotate_text for the number of observations in each category.
think about adding a vertical line for centromere, for the position means. Think about removing the extra Musc strains.

@@

Dom.male_MELT_long <- ggplot( long.biv_Dom.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Dom male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Dom.male_MELT_long

@@

Musc_male_MELT_long <- ggplot( long.biv_Musc.male, aes(x=PER.Foci.Pos, fill=strain) ) + geom_density(alpha=.2) + facet_wrap(~hand.foci.count, ncol=1)+ggtitle("Musc male long biv\nNormalized Foci Positions")+scale_fill_manual(values=colors_of_strains)
Musc_male_MELT_long

#there are v few observations in this dataset
```

<!-- Tired runnin plots for the long biv data set, I think there are too few observations. -->

@@

CO.Pos.boxplots_1CO <- ggplot(MELT_Curated_BivData.1CO, aes(x= strain, y=PER.Foci.Pos, fill=strain) ) + geom_boxplot(alpha=.2) + facet_wrap(sex~subsp, ncol=2, scales="free")+ggtitle("IFD across bivalents")+geom_jitter(alpha=.1)+scale_fill_manual(values=colors_of_strains)
CO.Pos.boxplots_1CO

@@

mouse.avs_4MM <- mouse.avs_4MM[!(is.na(mouse.avs_4MM$norm.first.foci.pos) | mouse.avs_4MM$norm.first.foci.pos==""), ]

#<- lme()
lme_MM_CO1_pos <- lme(norm.first.foci.pos ~ subsp * sex, data= mouse.avs_4MM, random=list(strain=pdDiag(~sex) ) )

as.data.frame( round(  summary(lme_MM_CO1_pos)$tTable, 6 ) )

@@

#library(lmerTest)
#anova(CO_position.MM)
#anova(CO_position.MM, verbose = TRUE)

#summary(CO_position.MM) #no p values,,
#these just give t and F stats, no pvalue

@@

#test high vs low Musc strains
#mouse.avs_4MM

@@

<!-- 
# old drafts

@@

#sis.co.ten.hist.poin99 <- ggplot((master_curated_BivData[1:50,]), aes(x=SisCoTen, y=chromosomeLength, color= #as.factor(hand.foci.count)  ))+
#  geom_point(aes(alpha=.5)) +facet_wrap(~ mouse)
```
-->
